An OpenCL Implementation of Wait-Free Sets

### 2015-07-03 Docs/paper.tex
doc changed

\begin{algorithm}
  \caption{snips next node if it is marked and could be snipped}
  \label{alg:snip}

  \begin{algorithmic}[1]
    \Require
        \Statex startRef:Reference to the starting node.
    \Ensure
       \Statex next node will be snipped.
    \Statex   
    \Function{snip}{startRef}
      \State pPRef = startRef:next;
      \State [nRef,pBits] = unpackRef(startRef:next);
      \If{(pBits = [0,0,0])}
        \State [nnRef,nBits] = unpackRef(pRef:next):
        \If{(nBits = [0,x,1])}
          \State status = CAS(startRef:next,pPRef,[nnRef,pBits]);
          \State
          \Return status;
        \Else
          \State
          \Return false;
        \EndIf
      \Else
        \State
        \Return false;
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

  \begin{algorithmic}[1]
    \Require
        \Statex startRef:Reference to the starting node.
    \Ensure
       \Statex next node will be replaced by new node.
    \Statex   
    \Function{replace}{startRef, newRef}
      \State pPRef = startRef:next;    
      \State [nRef,pBits] = unpackRef(startRef:next);
      \If{([pBits] = [0,0,0])}
        \State nPRef = nRef:next;    
        \State [nnRef,nBits] = unpackRef(nRef:next);
        \If{(nBits = [0,x,0])}
          \State status = CAS(nRef:next,nPRef,[nnRef,[0,1,0]]);
          \If{(status = false)}
            \State
            \Return false;
          \EndIf  
        \Else
          \State
          \Return false;
        \EndIf
        \State newRef:next = [nnRef,[0,0,0]];
        \State status = CAS(startRef:next,pPRef,[newRef,pBits]);
        \If{(status = false)}
          \State
          \Return false;
        \EndIf
      \Else
        \State
        \Return false;
      \EndIf
      \State
      \Return true;
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{cleans all the nodes that are logically deleted and could be physically deleted}
  \label{alg:clean}
  \begin{algorithmic}[1]
    \Require
        \Statex startRef:Reference to the starting node.
    \Ensure
       \Statex cleans all the logically deleted nodes.
    \Statex   
    \Function{clean}{startRef, nextRef\&}
      \State pRef  = startRef;
      \While{(done = 0)}
        \State nPRef = pRef:next;
        \State [nRef,prBit,pmBit] = unpackRef(nPRef);
        \If({prBit = 0})
          \State status = snip(pRef);
        \Else
          \State   
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\section{Proof of Correctness}
\begin{algorithm}
  \caption{Deletes all marked nodes that follow a given node}
  \label{alg:deleteMarkedNode}

  \begin{algorithmic}[1]
    \Require
        \Statex startRef:Reference to the starting node. Marked nodes
                that follow this are to be deleted. 
    \Ensure
       \Statex All marked nodes following {\it startRef} are deleted.
       \Statex Returns {\it true} if all marked nodes are deleted,
               {\it false} if node referenced by {\it startRef} gets deleted.

    \Statex   
    \Function{delMarkedNodes}{startRef}
      \State doneFlag = false;
      \While{(!doneFlag)}    
        \State [nextRef,startMark] := getUnmarkedRef(startRef);
        \If{(!startMark)}
          \State [nextNextRef,nextMark] := getUnmarkedRef(nextRef);
          \If{(nextMark)}
            \State status = CAS(startRef:next,[nextRef,startMark],[nextNextRef,startMark]);
            \If{(status = true)}
               \State freeNode(nextRef);
            \EndIf
          \Else
            \State
            \Return true;
          \EndIf
        \Else{}
          \State
          \Return false;  
        \EndIf
      \EndWhile        
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{theorem}
  A process running algorithm \ref{alg:deleteMarkedNode}, then it will either return next unmarked node after {\it startRef} in {\it nextRef}, or return {\it false}.
\end{theorem}
\begin{IEEEproof}
  The proof consists of the following parts.
  \begin{enumerate}
    \item The state of the linked list copied to internal variables {\it nextRef} and {\it nnRef} is valid as they are read simultaneously with {\it startMark} and {\it nextMark} respectively, and both these flags are checked.
    \item If this state remains valid, then CAS operation is successful.
    \item If this state is changed by any other process, this change is completely reflected in {\it startRef:next}.  
\end{enumerate} 
\end{IEEEproof}

\begin{algorithm}
  \caption{Check next node for key}
  \label{alg:check}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be checked.
       \Statex {\it prevRef}, reference to node next to which the key is to
               be checked.
    \Ensure
       \Statex Nodes referenced by {\it prevRef} and {\it prevRef:next} are
               unmarked. If they are marked, returns NODE\_DELETED and
               NODE\_NEXT respectively.
       \Statex If all values in {\it prevRef:next} $<$ {\it key}, moves
               {\it prevRef} to {\it prevRef:next}.          
       \Statex If values in {\it prevRef:next} $<$ {\it key}, checks for the key.
               Returns KEY\_FOUND or KEY\_NOT\_FOUND respectively.

    \Statex   
    \Function{checkKey}{key, prevRef\&}
       \State [nextRef,prevMark] := getUnmarkedPtr(prevRef);
       \If {(prevMark)}
         \State
         \Return {\small NODE\_DELETED}.
       \EndIf
       \If {(nextRef = 0)}
         \State
         \Return {\small KEY\_NOT\_FOUND}.
       \EndIf

       \State [nnRef,nextMark] := getUnmarkedPtr(nextRef);
       \If {(nextMark)}
         \State
         \Return {\small NODE\_NEXT}.
       \EndIf

       \State maxVal := simdMAX(nextRef:V[lid]);
       \If{maxVal $<$ key}
           \State prevRef := nextRef;
           \Return {\small NODE\_NEXT}.
       \Else
           \If{simdANY(nextRef:V[lid] = key)}
               \State
               \Return {\small KEY\_FOUND};
           \Else
               \State               
               \Return {\small KEY\_NOT\_FOUND};   
           \EndIf   
         \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find a key in the set}
  \label{alg:find}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be searched.
    \Ensure
       \Statex {\it prevRef}, such that next node either has key or key could
               be inserted to it.
       \Statex Returns {\it true} if key is found else returns {\it false}.

    \Function{find}{key, prevRef\&}
       \State lid := simdINDEX();
       \State startRef := hashFunction(key);
       \State prevRef  := startRef;
       \State doneFlag := false;
       \While{(!doneFlag)}
          \If{(lid = 1)}
             \State delStatus := delMarkedNode(prevRef);
          \EndIf
          \State simdBARRIER();
          \State delStatus = simdBROADCAST(delStatus,1);
          \If{delStatus = true}
             \State keyStatus := checkKey(key, prevRef);
             \If{(keyStatus = NODE\_DELETED)}
                \State prevRef := startRef;
             \ElsIf{(keyStatus = KEY\_FOUND)}
                \State 
                \Return true
             \ElsIf{(keyStatus = KEY\_NOT\_FOUND)}
                \State 
                \Return false
             \EndIf
          \Else
             \State prevRef := startRef;   
          \EndIf
       \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

>>>>

\begin{algorithm}
  \caption{Snips next node if it is marked and could be snipped}
  \label{alg:snip}

  \begin{algorithmic}[1]
    %\Require
    %    \Statex startRef:Reference to the starting node.
    %\Ensure
    %   \Statex  next node will be snipped.
    %\Statex   
    \Function{snip}{startRef}
      \State pPRef = startRef:next;
      \State [nRef,pBits] = unpackRef(startRef:next);
      \If{(pBits = [0,0,x])}
        \State [nnRef,nBits] = unpackRef(nRef:next):
        \If{(nBits = [0,x,1])}
          \State status = CAS(startRef:next,pPRef,[nnRef,pBits]);
          \If{(status = true)}
             \State FREE(nRef);
             \State 
             \Return {\it snipped};
          \Else
             \State
             \Return {\it failed};
          \EndIf
        \Else
          \State
          \Return {\it next\_unmarked};
        \EndIf
      \Else
        \State
        \Return {\it start\_invalid};
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

  \begin{algorithmic}[1]
    %\Require
    %    \Statex startRef:Reference to the starting node.
    %\Ensure
    %   \Statex next node will be replaced by new node.
    \Statex   
    \Function{replace}{startRef, newRef}
      \State pPRef = startRef:next;    
      \State [nRef,pBits] = unpackRef(startRef:next);
      \If{([pBits] = [0,0,0])}
        \State nPRef = nRef:next;    
        \State [nnRef,nBits] = unpackRef(nRef:next);
        \If{(nBits = [0,x,0])}
          \State status = CAS(nRef:next,nPRef,[nnRef,[0,1,0]]);
          \If{(status = false)}
            \State
            \Return {\it retain\_failed};
          \EndIf  
        \Else
          \State
          \Return {\it invalid\_nbits};
        \EndIf
        \State newRef:next = [nnRef,[0,0,0]];
        \State status = CAS(startRef:next,pPRef,[newRef,pBits]);
        \If{(status = true)}
          \State FREE(nRef);
          \State
          \Return {\it replaced}; 
        \Else
          \State
          \Return {\it failed};
        \EndIf
      \Else
        \State
        \Return {\it invalid\_pbits};
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{cleans all the nodes that are logically deleted and could be physically deleted}
  \label{alg:clean}
  \begin{algorithmic}[1]
    %\Require
    %    \Statex startRef:Reference to the starting node.
    %\Ensure
    %   \Statex cleans all the logically deleted nodes.
    \Statex   
    \Function{clean}{startRef, nextRef\&}
      \State pRef  = startRef;
      \State nextRef  = {\it null};
      \While{(true)}
        \State status = snip(pRef);
        \State [nRef,pBits] = unpackRef(pRef:next);
        \If{(status = {\it next\_unmarked})}
          \State nextRef = nRef;
          \State
          \Return status;
        \ElsIf{status = {\it start\_invalid}}
          \If{(pBits = [0,1,x])}
            \State pRef = nRef; 
          \Else
            \If{(pRef = startRef)}
              \State
              \Return {\it start\_invalid}
            \Else
              \State pRef = startRef;  
            \EndIf
          \EndIf
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{checks for valid window.}
  \label{alg:window}
  \begin{algorithmic}[1]
    %\Require
    %    \Statex startRef:Reference to the starting node.
    %\Ensure
    %   \Statex cleans all the logically deleted nodes.
    \Statex   
    \Function{window}{key, prevRef, nextRef, index\&}
      \State status = false;
      \If{(prevRef != {\it Null})}
        \State [nRef, pBits] = unpackRef(prevRef:next);
        \If{(pBits != [0,x,0])}
          \State 
          \Return {\it invalid\_prev\_bits}; ; 
        \EndIf
        \If{(nRef != nextRef)}
          \State 
          \Return {\it invalid\_next\_ref}; 
        \EndIf
        \State pMaxVal = simdMAX(prevRef:val);  
        \If{(key $<=$ pMaxVal)}
          \State 
          \Return {\it window\_not\_found};; 
        \EndIf
      \EndIf
      \State [nnRef, nBits] = unpackRef(nextRef:next);
      \If{(nBits != [0,x,0])}
        \State 
        \Return {\it invalid\_next\_bits}; 
      \EndIf
      \State pMaxVal = simdMAX(nextRef:val);  
      \If{(key $<=$ pMaxVal)}
        \State 
        \Return {\it window\_not\_found};
      \Else
        \If{(simdANY((key = nextRef:Val), index))}
          \State
          \Return {\it key\_found}
        \Else
          \State
          \Return {\it window\_found}
        \EndIf
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{finds a key and returns a window to it.}
  \label{alg:find}
  \begin{algorithmic}[1]
    \Require
        \Statex startRef:Reference to the starting node.
    \Ensure
       \Statex cleans all the logically deleted nodes.
    \Statex   
    \Function{find}{key, prevRef\&, nextRef\&, index\&}
      \State startRef = hash(key);
      \State prevRef = nextRef = {\it Null};
      \State index = 0;      
      \While(true)
        \State pRef = {\it Null};
        \State nRef = startRef;
        \State status = {\it window\_not\_found};
        \While{(status = {\it window\_not\_found})}
          \State status = WINDOW(key, pRef, nRef, index);
          \If{(status = {\it window\_not\_found})}
            \State pRef = nRef;
            \State status2 = CLEAN(pRef,nRef);
            \If{(status2 = {\it start\_invalid})}
              \State status = {\it window\_not\_found}
            \EndIf
          \EndIf
        \EndWhile
        \If{(status = {\it window\_found})}
          \State
          \Return status;
        \EndIf
        \If{(status = {\it key\_found})}
          \State
          \Return status;
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\section{Proof of Correctness}
